---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---
```{r}
library(ggplot2)
```

## Loading and preprocessing the data

load data
```{r}
myActivity <- read.csv("activity.csv", stringsAsFactors = FALSE)
str(myActivity)
```
convert date variable from char to date type
```{r}
myActivity$date <- as.Date(myActivity$date)
str(myActivity)
```
Check the data
```{r}
summary(myActivity)
```
## What is mean total number of steps taken per day?
1.What is mean total number of steps taken per day?
```{r}
myActivity_steps_per_day <- aggregate(steps ~ date, data = myActivity, FUN = sum, simplify = TRUE, na.rm = TRUE)
```
show histogram
```{r}
hist(myActivity_steps_per_day$steps, main = "Total number of steps taken per Day", xlab = "Total # Steps per Day")
```
show mean
```{r}
mean
```
show median
```{r}
median(myActivity_steps_per_day$steps)
```


## What is the average daily activity pattern?
2. What is the average daily activity pattern?
```{r}
myActivity_steps_mean <- aggregate(steps ~ interval, data = myActivity, FUN = mean, simplify = TRUE, na.rm = TRUE)
```
Plot result
```{r}
plot(myActivity_steps_mean$interval, myActivity_steps_mean$steps, type = "l", xlab = "5-Minutes-Intervals", ylab = "Average number of steps taken, averaged across all days ", main = "Average number of steps per interval")
```
Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
myMaxNosteps <-max(myActivity_steps_mean$steps)
myMaxInterval <- myActivity_steps_mean$interval[which(myActivity_steps_mean$steps == myMaxNosteps)]
myMaxInterval
myMaxNosteps
```

## Imputing missing values
3. Imputing missing values
Total amount of missing data
```{r}
sum(is.na(myActivity))
```
Create a new dataset from the original datasets with no missing data and using mean of the 5-minute interval
```{r}
myActivityNA<- subset(myActivity, !is.na(myActivity$steps))
myActivityNew <- myActivity
myNdx <- is.na(myActivityNew$steps)
myAvg <- tapply(myActivityNA$steps, myActivityNA$interval, mean, na.rm=TRUE, simplify=T)
myActivityNew$steps[myNdx] <- myAvg[as.character(myActivityNew$interval[myNdx])]
```
Total amount of missing data now (should be 0)
```{r}
sum(is.na(myActivityNew))
```
show histogram with imputed missing values
```{r}
myActivityStepsNew <- aggregate(steps ~ date, data = myActivityNew, FUN = sum, simplify = TRUE, na.rm = TRUE)
hist(myActivityStepsNew$steps, main = "Total number of steps taken per Day (missing values replaced)", xlab = "Total # Steps per Day")
```
show mean with imputed missing values
```{r}
mean(myActivityStepsNew$steps)
```
show median with imputed missing values
```{r}
median(myActivityStepsNew$steps)
```
-> The impact of imputed missing data very low
## Are there differences in activity patterns between weekdays and weekends?
4. Are there differences in activity patterns between weekdays and weekends?

Create a factor variable in the dataset with 2 levels: “weekday” and “weekend”
```{r}
myActivityNew$dayType <- ifelse(weekdays(as.Date(myActivityNew$date)) == "Samstag" | weekdays(as.Date(myActivityNew$date)) == "Sonntag", "weekend", "weekday")
myActivityNew$dayType <- factor(myActivityNew$dayType)
```
Aggregate a table showing mean steps for all intervals, acrlss week days and weekend days
```{r}
myStepIntervalDayType <- aggregate(steps ~ interval + dayType, data = myActivityNew, FUN = mean, simplify = TRUE, na.rm = TRUE)
```
add descriptive variables and plot
```{r}
names(myStepIntervalDayType) <- c("interval", "day_type", "mean_steps")
plot <- ggplot(myStepIntervalDayType, aes(interval, mean_steps))
plot + geom_line(color = "tan1") + facet_grid(day_type~.) + labs(x = "Intervals", y = "Average Steps", title = "Activity Patterns on Weekends/-days")
```
